---
output: md_document
---

<!-- CONTRIBUTING.md is generated from CONTRIBUTING.Rmd. Please edit that file -->

Contributing to a posterior database with R
============

First clone the posteriordb repository and install the posteriordb R package.

```{r, message=FALSE, eval=FALSE}
remotes::install_github("stan-dev/posteriordb-r")
```

Then loading the posteriordb R package and create a connection to the local posteriordb. If you intend to add stan-code, also load rstan.

```{r, message=FALSE, eval=FALSE}
library(posteriordb)
library(rstan)

# Set the environment variable PDB_PATH to you local posteriordb repo as
Sys.setenv(PDB_PATH = "[path to local posteriordb]")

# We setup a connection to the local posteriordb
pdbl <- pdb_local()
```

```{r, message=FALSE, eval=TRUE, echo=FALSE}
library(posteriordb)
library(rstan)

dotenv::load_dot_env(file = "../.env")

# We setup a connection to the local posteriordb
pdbl <- pdb_local()
```

## Add Data

The next step is to add and write down information about the data you want to add. Below is a test example using data from the eight schools example.


```{r}
x <- list(name = "test_eight_schools_data",
          keywords = c("test_data"),
          title = "A Test Data for the Eight Schools Model",
          description = "The data contain data from eight schools on SAT scores.",
          urls = "https://cran.r-project.org/web/packages/rstan/vignettes/rstan.html",
          references = "rubin1981estimation",
          added_date = Sys.Date(),
          added_by = "Stanislaw Ulam")

# Create the data info object
di <- pdb_data_info(x)

# Access the data
file_path <- system.file("test_files/eight_schools.R", package = "posteriordb")
# You can check the file with: file.edit(file_path)
source(file_path)

# Create the data object
dat <- pdb_data(eight_schools, info = di)
```


We can now add the data object to the database (or remove it). The function ```write_pdb()``` will write and zip the data.

```{r}
write_pdb(dat, pdbl)
```

If we want to remove the data object, we can simply use

```{r, eval=FALSE}
remove_pdb(dat, pdbl)
```

Note that we now added a reference `rubin1981estimation` that already exists in the bibliography. If a new reference is added, it also needs to be added in the bibtex bibliography.

## Add Model

Similarly, we can add stan-files and model information as follows (this is a test case using the eight school model): 

```{r}
x <- list(name = "test_eight_schools_model",
          keywords = c("test_model", "hiearchical"),
          title = "Test Non-Centered Model for Eight Schools",
          description = "An hiearchical model, non-centered parametrisation.",
          urls = c("https://cran.r-project.org/web/packages/rstan/vignettes/rstan.html"),
          framework = "stan",
          references = NULL,
          added_by = "Stanislaw Ulam",
          added_date = Sys.Date())
mi <- pdb_model_info(x)

# Read in Stan model and compile the model
file_path <- system.file("test_files/eight_schools_noncentered.stan", package = "posteriordb")
smc <- readLines(file_path)
sm <- rstan::stan_model(model_code = smc)
mc <- model_code(sm, info = mi)

# Write the model to the database
write_pdb(mc, pdbl)
```

Similarly we remove the object with
```{r, eval=FALSE}
remove_pdb(mc, pdbl)
```


## Add Posterior Object

```{r}
x <- list(pdb_model_code = mc,
          pdb_data = dat,
          keywords = "posterior_keywords",
          urls = "posterior_urls",
          references = "rubin1981estimation",
          dimensions = list("dimensions" = 2, "dim" = 3),
          reference_posterior_name = NULL,
          added_date = Sys.Date(),
          added_by = "Stanislaw Ulam")
po <- pdb_posterior(x, pdbl)

# We write to the database as done previously
write_pdb(po, pdbl)
```

And to remove the posterior, we simply use
```{r, eval=FALSE}
remove_pdb(po, pdbl)
```

Note that we need to add the model code and data first, before adding the posterior.

Finally, we want to check that everything is in order with the posterior. We do this as follows:

```{r}
check_pdb_posterior(po)
```

If the posterior passes all checks, it can be added to the posteriordb, so it is just to open a Pull Request with the proposed posterior.


## Add Posterior Reference Draws

If possible, we would like to supply posterior reference draws, i.e., draws of excellent quality from the posterior. The [REFERENCE_POSTERIOR_DEFINITION.md](https://github.com/MansMeg/posteriordb/blob/master/doc/REFERENCE_POSTERIOR_DEFINITION.md) contain details on quality criteria for reference posteriors.

```{r}
pdbl <- pdb_local()
po <- posterior("test_eight_schools_data-test_eight_schools_model", pdbl)

# Setup reference posterior info ----
x <- list(name = posterior_name(po),
          inference = list(method = "stan_sampling",
                           method_arguments = list(chains = 10,
                                                   iter = 20000,
                                                   warmup = 10000,
                                                   thin = 10,
                                                   seed = 4711,
                                                   control = list(adapt_delta = 0.92))),
          diagnostics = NULL,
          checks_made = NULL,
          comments = "This is a test reference posterior",
          added_by = "Stanislaw Ulam",
          added_date = Sys.Date(),
          versions = NULL)

# Create a reference posterior draws info object
rpi <- pdb_reference_posterior_draws_info(x)
```

The reference posterior draws info contain all information to compute the posterior using rstan. We then check that the reference posterior criteria are fulfilled and add checked diagnostics to the object.

```{r, eval=FALSE}
# Compute the reference posterior
rp <- compute_reference_posterior_draws(rpi, pdbl)
rp <- check_reference_posterior_draws(x = rp)
```

We can now write the reference posterior draws to the posteriordb.

```{r, eval = FALSE}
write_pdb(rp, pdbl, overwrite = TRUE)

# We can again check the posterior with
check_pdb_posterior(po)
```

We can also remove the reference posterior object with
```{r, eval=FALSE}
remove_pdb(rp, pdbl)
```

```{r cleanup, eval=TRUE, echo=FALSE}
# Cleanup
invisible(remove_pdb(mc, pdbl))
invisible(remove_pdb(dat, pdbl))
invisible(remove_pdb(po, pdbl))
```


